# Functions
```{r}
str_entropy <- function(s) {
    f <- table(strsplit(s, "")) / nchar(s)
    print(f)
    h <- 0
    for (i in f) {
        if (i != 0) {
            h <- h + i * log2(i)
        }
    }
    return(-h)
}

str_entropy2 <- function(s, k_len) {
    #----- contingency table for probability -----#
    nb_kmer <- nchar(s) - k_len + 1
    kmer_list <- c()
    for (pstart in 1 : nb_kmer) {
        kmer_list <- c(kmer_list, substr(s, pstart, pstart + k_len - 1))
    }
    f <- table(kmer_list) / sum(table(kmer_list))
    print(f) # print the contingency table for check #
    #----- sum all probabilities -----#
    h <- 0
    for (i in f) {
        if (i != 0) {
            h <- h + i * log2(i)
        }
    }
    return(-h)
}
```

# A Sampe Case
For a sequence of repetition of (ATC) motif with an supplementary A at the end, the entropie does not decrease by taking 3-mers instead of single nucleotide for elementary motifs.

```{r}
print(paste("Entropie:", round(str_entropy("ATCATCATCATCATCATCATCATCATCATCA", 1), 6)))
```

```{r}
print(paste("Entropie:", round(str_entropy("ATCATCATCATCATCATCATCATCATCATCA", 2), 6)))
```

```{r}
print(paste("Entropie:", round(str_entropy("ATCATCATCATCATCATCATCATCATCATCA", 3), 6)))
```

# An Actual k-mer from Monkeypox CDS Sequence
For an actual k-mer from the problematic Monkeypox CDS, the entropie even increases when changing elementary motif from single nucleotide to 3-mer.

```{r}
print(paste("Entropie:", round(str_entropy("GATGATGATGATGATATCATAGATGATGATG", 1), 6)))
```

```{r}
print(paste("Entropie:", round(str_entropy("GATGATGATGATGATATCATAGATGATGATG", 2), 6)))
```

```{r}
print(paste("Entropie:", round(str_entropy("GATGATGATGATGATATCATAGATGATGATG", 3), 6)))
```